Modeling the near-wake of a vertical-axis cross-flow turbine with 2-D and 3-D RANS

### 2016-10-02 reviews/reviewer2-response.md

Since the chosen SA time step was about half that of the SST, the SA model did
take longer to solve, though they were still the same order of magnitude. We
decided it wouldn't be useful to readers to give very specific metrics of solver
time since there could be many intervening variables, e.g., hardware.

>Did you prove that the meshing of the outer walls didn't affect the results.

We did not, but if we inspect the velocity fields, which don't extend out all
the way to the walls, gradients are very small around the outer region,
indicating the effects should be negligible.

>why approximately [rotor performance periodicity]?

Since this is an unsteady simulation, there are interactions between blade wakes
that are not perfectly periodic, so there are some minor differences between
rotor revolutions. Note that this is consistent with the experimental results.

@@

>Why blank? [viscous transport in Figure 12]

These terms aren't zero, but are extremely small.

>Not sure if this is true - DMS models have tip correction factors - see reference #2?

We could not find any instances in this reference for DMS model tip correction
factors for VAWTs. However, if the reviewer has a specific correction in mind we
would be happy to note this in the manuscript.

### 2016-10-02 reviews/reviewer2-response.md

These terms aren't zero, but are extremely small.

>Not sure if this is true - DMS models have tip correction factors - see reference #2?

We could not find any instances in this reference for DMS model tip correction
factors for VAWTs. However, if the reviewer has a specific correction in mind we
would be happy to note this in the manuscript.

>Can you show differences between these graphically, i.e subtract B from A and C from A to show the change - this would make it a lot easier to see differences

We did try computing "difference" plots, but these were a bit noisy and were
difficult to interpret.

>Wouldnt this have to be a number of complete revolutions?

It virtually is--6.03 revolutions technically, but the additional 0.03
revolutions should not have any appreciable effect on the statistics.

### 2016-10-02 paper.tex
Clarify which formulas indicate which recovery terms

To get an overall idea of the wake recovery predicted by each model, and
therefore its potential effectivness in evaluating array spacing, the streamwise
component of the Navier--Stokes equation was rearranged to isolate $\partial U /
\partial x$---following Bachant and Wosnik~\cite{Bachant2015-JoT}---and compute
each term at $X/D = 1$ to compare with the experimental results.

@@

Normalized weighted averages for each recovery term at $x/D=1$ are computed and
multiplied by the cross-sectional area of the measurement plane, or the channel
width in the 2-D cases. Results are shown in a bar chart in
Figure~\ref{fig:br-cfd-recovery}. Consistent with the relatively large Reynolds
number regime, viscous transport is essentially negligible compared with other
mechanisms. Cross-stream advection---or the tendency of streamlines to diverge
and reduce the streamwise momentum---produces a negative effect for all cases,
though the 3-D SST model predicts significantly lower values. Vertical advection
is by definition zero for the 2-D cases. The 3-D cases show varying
results---with the SST model overpredicting and SA underpredicting the vertical
velocity's effect on replenishing streamwise momentum. Nevertheless, both 3-D
models were able to qualitatively resolve this mean vertical advection caused by
blade tip vortex shedding, which will be crucial to predicting overall wake
recovery and the performance of closely spaced turbines.

Turbulent transport and streamwise pressure gradient terms show the largest
discrepancy between results. The 3-D SST case, despite doing a good job
predicting turbulence kinetic energy, significantly overpredicts the turbulent
transport term, while other CFD cases are comparable with the experiment. This
seems to be balanced by a large adverse pressure gradient, which is also present
to a smaller degree in the 3-D SA case. Interestingly, in contrast, both 2-D CFD
cases create a wake where the pressure gradient is acting to slightly accelerate
the flow at $x/D=1$. Unfortunately, pressure data were not acquired from the
experiment, though results from the 3-D delayed detached eddy simulation (DDES)
of Boudreau and Dumas \cite{Boudreau2015} concur with the adverse pressure
gradient condition.

>>>>

To get an overall idea of the wake recovery predicted by each model, and
therefore its potential effectivness in evaluating array spacing, the streamwise
component of the Navier--Stokes equation was rearranged to isolate the
streamwise partial derivative of the streamwise mean velocity $\partial U /
\partial x$---following Bachant and Wosnik~\cite{Bachant2015-JoT}---and compute
each term at $x/D = 1$ to compare with the experimental results.

@@

Normalized weighted averages for each recovery term at $x/D=1$ are computed and
multiplied by the cross-sectional area of the measurement plane, or the channel
width in the 2-D cases. Results are shown in a bar chart in
Figure~\ref{fig:br-cfd-recovery}. Consistent with the relatively large Reynolds
number regime, viscous transport is essentially negligible compared with other
mechanisms. Cross-stream advection $-V \frac{\partial U}{\partial y}$; or the
tendency of streamlines to diverge and reduce the streamwise momentum---produces
a negative effect for all cases, though the 3-D SST model predicts significantly
lower values. Mean vertical advection $-W \frac{\partial U}{\partial z}$ is by
definition zero for the 2-D cases. The 3-D cases show varying results---with the
SST model overpredicting and SA underpredicting the vertical velocity's effect
on replenishing streamwise momentum. Nevertheless, both 3-D models were able to
qualitatively resolve this mean vertical advection caused by blade tip vortex
shedding, which will be crucial to predicting overall wake recovery and the
performance of closely spaced turbines.

Turbulent transport and mean streamwise pressure gradient ($-\partial P \partial
x$) terms show the largest discrepancy between results. The 3-D SST case,
despite doing a good job predicting turbulence kinetic energy, significantly
overpredicts the turbulent transport term, while other CFD cases are comparable
with the experiment. This seems to be balanced by a large adverse pressure
gradient, which is also present to a smaller degree in the 3-D SA case.
Interestingly, in contrast, both 2-D CFD cases create a wake where the pressure
gradient is acting to slightly accelerate the flow at $x/D=1$. Unfortunately,
pressure data were not acquired from the experiment, though results from the 3-D
delayed detached eddy simulation (DDES) of Boudreau and Dumas
\cite{Boudreau2015} concur with the adverse pressure gradient condition.
